---
title: "fire2"
output: html_document
---


Data: https://www.kaggle.com/carlosparadis/fires-from-space-australia-and-new-zeland?select=fire_archive_M6_96619.csv

https://earthdata.nasa.gov/earth-observation-data/near-real-time/firms/c6-mcd14dl 

#FEATURES

Latitude: Center of 1km fire pixel but not necessarily the actual location of the fire as one or more fires can be detected within the 1km pixel.

Longitude: Center of 1km fire pixel but not necessarily the actual location of the fire as one or more fires can be detected within the 1km pixel.

Brightness: Brightness temperature 21 (Kelvin): Channel 21/22 brightness temperature of the fire pixel measured in Kelvin.

Scan: Along Scan pixel size: The algorithm produces 1km fire pixels but MODIS pixels get bigger toward the edge of scan. Scan and track reflect actual pixel size.

track: Along Track pixel size: The algorithm produces 1km fire pixels but MODIS pixels get bigger toward the edge of scan. Scan and track reflect actual pixel size.

acq_date: Acquisition Date: Date of MODIS acquisition.

acq_time: Acquisition Time: Time of acquisition/overpass of the satellite (in UTC).

Satellite: A = Aqua and T = Terra.

Instrument: Constant value for MODIS. 

Confidence (0-100%): This value is based on a collection of intermediate algorithm quantities used in the detection process. It is intended to help users gauge the quality of individual hotspot/fire pixels. Confidence estimates range between 0 and 100% and are assigned one of the three fire classes (low-confidence fire, nominal-confidence fire, or high-confidence fire).

Version (Collection and source): Version identifies the collection (e.g. MODIS Collection 6) and source of data processing: Near Real-Time (NRT suffix added to collection) or Standard Processing (collection only).
"6.0NRT" - Collection 6 NRT processing. "6.1NRT" - Collection 61 NRT processing
"6.0" - Collection 6 Standard processing. "6.1" - Collection 61 Standard processing

Brightness temperature 31 (Kelvin): Channel 31 brightness temperature of the fire pixel measured in Kelvin.

Fire Radiative Power (MW - megawatts): Depicts the pixel-integrated fire radiative power in MW (megawatts).

Type*
0 = presumed vegetation fire
1 = active volcano
2 = other static land source
3 = offshore
*This attribute is only available for MCD14ML (standard quality) data 



```{r, echo=FALSE,  message=FALSE }
library(sf)
library(mapview)
library(ggplot2)
library(readr)
library(tidymodels)
library(dplyr)
library(caret)

```


```{r, echo=FALSE}
fire_dataset <- read.csv("./fire_archive_M6_96619.csv")
```


# MAP OF THE FIRES IN AUSTRALIA
```{r, echo=FALSE}

plot <- st_as_sf(fire_dataset, coords = c("longitude", "latitude"),  crs = 4326)
mapview(plot)

```


# PLOT THE OUTPUT VARIABLE

```{r, echo=FALSE}

fire_dataset %>%
  ggplot( aes(x=frp)) +
    geom_density(fill="brown3", color="brown3", alpha=0.8)


mean_frp <- mean(fire_dataset$frp)
median_frp <- median(fire_dataset$frp)
min_frp <- min(fire_dataset$frp)
max_frp <-  max(fire_dataset$frp)

library(gt)
brg_tbl <- tibble(
    mean_frp,
    median_frp,
    min_frp,
    max_frp
  ) 

gt_tbl <- gt(data = brg_tbl)
gt_tbl
```

# Transform as date

```{r, echo=FALSE}

fire <- fire_dataset %>% mutate(acq_date = as.Date(acq_date, format = "%Y-%m-%d"),
                       satellite = as.factor(satellite),
                       instrument = as.factor(instrument),
                       daynight = as.factor(daynight)
                       )

```




# no missing values

```{r, echo=FALSE}
library(naniar)
vis_miss(fire)
```



# correlation plot

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(PerformanceAnalytics)
mydata <- fire[, c(-6,-8,-9, -14)]
chart.Correlation(mydata, histogram=TRUE, pch=19)
#http://www.sthda.com/french/wiki/matrice-de-correlation-guide-simple-pour-analyser-formater-et-visualiser
```
The variables scan and Track are highly correlated with 0,98.
bright_t31 and brightness are strong correlated. 
frp and brightness are highly correlated to brightness.


# PLOT BRIGHTNESS ACCORDING TO ACQUISITION DATE DISTINCTION BETWEEN DAY AND NIGHT
```{r}
ggplot(fire, aes(x = acq_date, y = brightness))+
  geom_point(aes(color = daynight, shape = daynight)) +
   scale_color_manual(values=c("darkorange2", "aquamarine4"))
  
```

# CHANGE TIME??


```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(fire, aes(x = acq_time, y = brightness))+
  geom_point(aes(color = daynight, shape = daynight)) +
   scale_color_manual(values=c("firebrick1", "darkcyan")) +
   facet_wrap(~ satellite)

```



# PLOT BRIGHTNESS ACCORDING DAY OR NIGHT FACTOR, COLORS ARE REPRESENTING THE DIFFERENT SATELLITES

```{r}
library(gganimate)
p_1 <- ggplot(
  fire, 
  aes(x = scan, y = brightness , colour = satellite)
  ) +
  geom_point(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(0, 1)) +
  scale_x_log10() +
  labs(x = "Day or night ", y = "brightness")

p_1  + transition_time(acq_date) + ggtitle('Eathquake magnitude type defined per Magnitude and per year ',
          subtitle = 'acq_date: {frame_time}')

```



# fACED WRAP ACCORDING TYPE 

```{r}

ggplot(fire, aes(x = scan, y = brightness))+
  geom_point(aes(color = daynight, shape = daynight)) +
  scale_color_manual(values=c("brown1", "darkorchid4")) +
   facet_wrap(~ type)

```


# PLOT OFFSHORE TYPE WILDFIRES

```{r, echo=FALSE}
fire_offshore <- filter(fire, type == 3)
plots_offshore <- st_as_sf(fire_offshore, coords = c("longitude", "latitude"),  crs = 4326)
mapview(plots_offshore)

```


# PLOT LAND TYPE WILDFIRE
```{r, echo=FALSE}
fire_land <- filter(fire, type == 2)
plots_land <- st_as_sf(fire_land, coords = c("longitude", "latitude"),  crs = 4326)
mapview(plots_land)

```


#PLOT Brightness temperature 31 (Kelvin) according to track

```{r}
p_2 <- ggplot(
  fire, 
  aes(x = track, y = frp , colour = type)
  ) +
  geom_point(show.legend = TRUE, alpha = 0.7) +
  scale_color_viridis_b (option = "magma") +
  scale_size(range = c(0, 1)) +
  scale_x_log10() +
  labs( x = 'track', y = 'bright_t31')



p_2 + transition_time(brightness) + ggtitle(' Brightness temperature 31 (Kelvin) according to track',
          subtitle = 'brightness: {frame_time}')


```



```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(plotly)
fig <- plot_ly(fire, x = ~acq_date, y = ~frp, z = ~scan, 
               color = ~satellite, # define color on param
               colors = c('cornflowerblue', 'goldenrod1'))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Date'),
                     yaxis = list(title = 'frp'),
                     zaxis = list(title = 'scan')))

fig

#https://jtr13.github.io/cc21/d-plot-in-r.html

```







# remove instrument, version (just one value) and confidence
```{r}
fire_modelling = select(fire, -9, -10, -11)
str(fire_modelling)
```


# split dataset using caret library
```{r}
uin = 123456789
set.seed(uin)
fi_split = initial_split(fire_modelling, p = 0.6, list = FALSE)
fi_trn = fi_split %>% training()
fi_tst = fi_split %>% testing()
```


# grid for tuning 
```{r}
gbm_fi_grid = expand_grid(interaction.depth = c(1, 2),
                       n.trees = c(500, 1000, 1500),
                       shrinkage = c(0.001, 0.01, 0.1),
                       n.minobsinnode = 10)
```

# cross validation 
```{r}
set.seed(uin)
cv_5 = trainControl(method = "cv", number = 5)

set.seed(uin)
fit_control_oob <- trainControl(
                           method = "oob",
                           number = 5)
```



# modelling 
```{r}
#xgboost
set.seed(uin)
fi_gbm = train(frp ~ . , data = fi_trn,
                method = "gbm",
                trControl = cv_5,
                verbose = FALSE,
                tuneGrid = gbm_fi_grid)
```




# performance and testing 
```{r}
df_perf_fi_gbm <- getTrainPerf(fi_gbm) %>% glimpse()
fi_gbm_rmse_train <- df_perf_fi_gbm$TrainRMSE  
fi_gbm_pred <- fi_gbm %>% predict(fi_tst)
fi_gbm_rmse_test <- RMSE(fi_gbm_pred, fi_tst$frp)
```










